---
title: "Regression Model Course Project"
author: "Sachin Sharma"
date: "10/7/2021"
output:
  html_document: default
  pdf_document: default
  always_allow_html: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)


```


\textcolor{red}{\Huge \textbf{Brief Summary}}

\textcolor{blue}{\LARGE In this report, we will examine the \textbf{mtcars} data set and explore how miles per gallon (MPG) is affected by different variables. As per the requiremnt of the project, we will answer the following two questions:}

\begin{enumerate}

\item Is an automatic or manual transmission better for MPG, and 

\item  Quantify the MPG difference between automatic and manual transmissions.


\end{enumerate}




\textcolor{red}{\Huge \textbf{Exploratory Data Analysis}}


\textcolor{red}{\LARGE STEP :1 - \textbf{Importing Libraries}}

```{r}
library(tidyverse)
library(ggplot2)
library(naniar)
library(dplyr)
library(datasets)
library(tinytex)
library(DT)


```


\textcolor{red}{\Huge \textbf{Reading Data}}

# Here we can use datatable, which is very useful to filter the data and see the quick results. However, I have separated the data in my next steps using separate codes. 

```{r}

data("mtcars")

datatable(mtcars,rownames = FALSE,filter = "top")

head(mtcars)



```

# Transform certain variables into factors

```{r}

mtcars$cyl  <- factor(mtcars$cyl)
mtcars$am   <- factor(mtcars$am,labels=c("Automatic","Manual"))
mtcars$vs   <- factor(mtcars$vs)
mtcars$gear <- factor(mtcars$gear)
mtcars$carb <- factor(mtcars$carb)







```


```{r}

boxplot(mpg ~ am, data = mtcars, col = (c("purple","red")), ylab = "Miles Per Gallon", xlab = "Type of Transmission", main = "MPG Vs AM")


```




\textcolor{red}{\LARGE \textbf{Regression Analysis}}

\textbf{\Large  With the help of plot, We’ve visually seen that automatic is better for MPG, but we will now quantify his difference}


```{r}

aggregate(mpg~am, data = mtcars, mean)






```

## Difference of MPG between Automatic and Manual 

```{r}
24.39231 - 17.14737

```


## Therefore, we can see that the Manual cars have an MPG of 7.245 (approx.) more than automatic cars 


## We  can now use a t-test here 


<h3 style="color:red" >What is t-test ? </h3>


<h3  style="color:blue">The t-test assesses whether the means of two groups are statistically different from each other.This analysis is appropriate whenever you want to compare the means of two groups</h3>



```{r}
automatic_car <- mtcars[mtcars$am == "Automatic",]
manual_car <- mtcars[mtcars$am == "Manual",]
t.test(automatic_car$mpg, manual_car$mpg)


```

# We can see that the p-value is 0.001374, thus we can state this is a significant difference. Now to quantify this, we can use the following code : 



```{r}

model_1  <- lm(mpg ~ am, data = mtcars)
summary(model_1)




```
<h3 style="color:blue"   > The above data shows us that the average MPG for automatic is 17.1 MPG, while manual is 7.2 MPG higher. The $R^2$ value is 0.36 which states that, this model only explains us 36% of the variance. As a result, here we require to build a multivariate linear regression.</h3>


## Lets see with the help of corrplot , to check the correlation among the variables with mpg.

## Before plotting the corrplot, we will check the structure of the data ;  

```{r}


df_1 <- subset(mtcars, select = c(mpg,cyl,disp,hp,drat,wt,qsec,vs))

head(df_1)



str(df_1)



```

# Here we can see that, cyl and vs columns are in factor, we will now convert this into numeric to plot corrplot and check the correlation. 


```{r}

df_1$cyl <- as.character(df_1$cyl) 

df_1$cyl <- as.numeric(df_1$cyl)

df_1$vs <- as.character(df_1$vs)

df_1$vs <- as.numeric(df_1$vs)


# Now we can check the structure of the data again 
str(df_1)

```

# Now we can see that all the columns are in numeric, now we can plot wit the help of ggcorrplot and corrplot to check the correlation : 




```{r}
library(ggcorrplot)

r <- cor(df_1)

ggcorrplot(r,method = "circle", type = c("upper"), legend.title = "Corrplot MTCARS")

```
```{r}

library(corrplot)
r <- cor(df_1)


corrplot(r, method = "circle")

```

<h3 style="color:blue" > With the help of above two plots, we can easily say that cyl,disp, hp and wt have strong correlation  with mpg </h3>

<h3 style="color:blue" > We build a new model using these variables and compare them to the initial model with the anova function.</h3>

```{r}
model_2  <- lm(mpg~am + cyl + disp + hp + wt, data = mtcars)
anova(model_1, model_2)

```

<h3 style="color:blue" > Here we can see that the result of p-value is 8.637e-08, and hence we can say that  our model_2 is significantly better than our model_1 which is a simple model.</h3>


<h3 style="color:blue" > We can plot the graph to check the residuals for non - normality and see whether they are normally distributed or not. </h3>



```{r}
par(mfrow = c(2,2))
plot(model_2)
```

# Now we will check the summary of our model_2 


```{r}

summary(model_2)
```

<h3 style="color:blue" >With the help of the above summary, we can say that model explain that there is a variance of 86.64% and as a result variables like cyl, disp, hp, wt did affect the correlation between mpg and am. </h3>


<h3 style="color:blue" > Hence, we can say the difference between automatic and manual transmissions is 1.81 MPG </h3>




